Artificial Intelligence for the Automated Synthesis and Validation of Programs

### 2018-09-11 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updated evaluation project proposal

  Our current work on this research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and recently, it is been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for research in AI.

@@

\subsection{Specific objectives}
\label{sec:objectivos}
We characterize the programs objects of study in this project by these tree dimensions:
\begin{enumerate}
\item Number of {\em program lines}.
\item Number and domain of the observable {\em program variables}.
\item Size of the available {\em instruction set}.
\end{enumerate}

To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ is generated using three program lines, two observable Boolean variables ($x=N$ and $y=N$), and an instruction set that comprises four instructions. Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em preparing an orange juice}, is generated using five program lines, the observable Boolean variable $ready(juice)$ that holds when the orange juice is recognized to be ready, and an instruction set that comprises four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$).

The specific objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of TDD programming tasks coming from the two kinds of domains described in the previous subsection. In further detail, the two specific objectives of this project are:
\begin{enumerate}
\item To study the performance of our Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}. 
\item To study the performance of our Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}. 
\end{enumerate}

Despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}. With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.

@@

We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that (1), takes as input a set of test cases that specify the TDD programming task to solve and (2), outputs a program source code that passes these tests with a bug-free guarantee.

In the particular case that the {\em program synthesizer} receives an additional input specifying the program source code, it outputs a validation certificate guaranteeing that the input program passes the given test cases. Figure~\ref{fig:gantt} details the proposed 24-month timeline for the project. A deliverable is provided at the end of each task.

@@

\subsection{Workload}

{\em Tasks ({\bf T1-T3})} will be developed by the three mentioned members of the research group: Sergio Jiménez, Eva Onaindia and Diego Aineto.

In addition, we plan to hire a master student for 6 months to assist in the completion of two tasks, {\bf T3(a,i)} and {\bf T3(a,iv)}. The detailed responsability of the student will be:
\begin{enumerate}
\item Executing the scripts of the system architecture.
\item Collecting the output data.
\item Presenting the output data in a easy-readable format.
\end{enumerate}

>>>>

  Our current work on this research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and it is been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for research in AI.

@@

\subsection{Specific objectives}
\label{sec:objectivos}
We characterize the programs objects of study in this project by these three dimensions:
\begin{itemize}
\item Number of {\em program lines}.
\item Number and domain of the observable {\em program variables}.
\item Size of the available {\em instruction set}.
\end{itemize}

To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ is generated using: three program lines, two observable Boolean variables ($x=N$ and $y=N$), and an instruction set that comprises four instructions. Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em preparing an orange juice}, is generated using: five program lines, the observable Boolean variable $ready(juice)$ that holds when the orange juice is recognized to be ready, and an instruction set that comprises four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$).

The objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of TDD programming tasks that come from the two kinds of domains described in the previous subsection. In further detail, the specific objective can be itemized into two specific objectives:
\begin{enumerate}
\item To study the performance of our Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}. 
\item To study the performance of our Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}. 
\end{enumerate}

Note that despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}. With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.

@@

We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that:
\begin{enumerate}
\item Takes as input a set of test cases that specify the TDD programming task to solve and outputs a program source code that passes these tests with a bug-free guarantee.
\item The {\em program synthesizer} must also work in {\em validation a mode}. In this particular setting, the {\em program synthesizer} receives an additional input specifying the program source code and it outputs a {\em validation certificate} guaranteeing that the input program passes (or fails) at the given test cases. 
\end{enumerate}

Figure~\ref{fig:gantt} details the proposed 24-month timeline for the project. A deliverable is provided at the end of each task.

@@
